This work would not exist without Professor Dustin Richmond, my advisor.
I am grateful for his support, feedback, and willingness to spend
countless hours poring over timing data with me. Beyond the work, he
made me into someone capable of doing it.

I am also grateful to the rest of my committee. To Professor Ryan
Kastner, who kept me focused on the larger picture. And to Professors
Jason Eshraghian and Matthew Guthaus for the time and attention they
have given to this committee.

I thank Dr. Alric Althoff, whose feedback put me on the path to solving
the error-correcting code parity recovery problem and whose
collaboration helped to shape the leakage channel in the BREW-RC work. I
also thank Professor David Kohlbrenner, whose suggestion to cast
reverse-engineering problems as SAT formulations inspired the SAT
construction used to recover error-correcting code.

My path into research, hardware security, and emerging memories began
while working in the Nano Electronics and Computing Research Lab with
Professor Hamid Mahmoodi at San Francisco State University. The ideas
and questions he put in front of me then are in many ways the ones I'm
still asking.

Finally, to my lab mates Ella Lehavi, Aditya Bedekar, Eddy Abban, and
Ethan Sifferman, and to the many undergraduate and Master's students who
have engaged with my work along the way: thank you for the
conversations, the questions that sent me back to the whiteboard, and
the patience to engage with ideas before they had taken shape.

\section{Publications}

This dissertation is based off of three works. The following
first-author works:
\begin{itemize}
	\item \fullcite{sheaves2026brewrc}
	\item \fullcite{sheaves2026ecc}~\textbf{[Under Review]}
\end{itemize}
and the following co-first author work:
\begin{itemize}
	\item \fullcite{drewes2024turn}
\end{itemize}
For the co-first authored work, this dissertation contains only my exact
contribution.

\section{Funding Acknowledgement}

This material is based upon work supported by the National Science
Foundation under Award No. 2418268. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of
the author and do not necessarily reflect the views of the National
Science Foundation.

