\section{Research Objectives and Contributions}
\label{sec:objectives}

This dissertation pursues four research objectives:

\begin{enumerate}
	\item \textbf{Broaden FPGA Pentimenti measurement across vendors.} We develop
	      a constrained PLL configuration search that achieves the fine-grained
	      timing resolution required for Pentimenti extraction on FPGAs that do
	      not support fine-grained phase adjustments. We demonstrate these
	      techniques on an Intel Cyclone V FPGA, closing the prior dependence on
	      Xilinx-specific high-resolution phase shifting capabilities.
	\item \textbf{Characterize commercial ReRAM write latency.} We identify and
	      experimentally validate the dominant structural factors that affect
	      write latency in commercial ReRAM crossbar memories, including crossbar
	      architecture, embedded ECCs, and device aging.
	\item \textbf{Recover hidden device architecture non-invasively.} We develop a
	      technique to recover the bit-exact ECC parity submatrix from commercial
	      ReRAM products using only write-access timing, without physical probing
	      or device modification.
	\item \textbf{Apply structural knowledge to entropy characterization.} We
	      demonstrate that ECC-aware analysis is necessary for rigorous
	      characterization of stochastic write latency, and construct entropy
	      sources whose min-entropy can be evaluated under NIST SP 800-90B
	      requirements. In characterizing the entropy of write latency under
	      various wear-levels, we identify the connection between write entropy
	      and ReRAM filament degradation, providing a concrete demonstration that
	      our measurement techniques are sensitive to conductive filament defect
	      dynamics.
\end{enumerate}

This dissertation makes the following contributions:

\paragraph{Cross-Vendor FPGA Pentimenti Sensor Implementation.}
We present a vendor-agnostic implementation of FPGA Pentimenti sensors
that closes the prior dependence on Xilinx-specific high-resolution PLL
phase shifting capabilities. By reformulating fine-grained sensor tuning
as a constrained PLL configuration search over valid PLL configurations
we achieve the timing resolution required for Pentimenti extraction on
Intel Cyclone V FPGAs. This broadens the FPGA platforms on which
Pentimenti analysis is practical.

\paragraph{Timing Side-Channel Characterization.}
We present the first systematic characterization of write latency in
commercial ReRAM crossbar memories, identifying an array-level timing
property called \emph{polarity serialization} and embedded ECC activity
as the dominant sources of structured timing variance. We introduce the
directional Hamming distance (DHD) as a metric for classifying write
transitions and demonstrate that ECC parity bit activity is large enough
to obscure the stochastic device physics that prior security and entropy
analyses assumed they were measuring. These findings are validated on
four samples each of two commercial RAMXeed ReRAM products.

\paragraph{BREW-RC: Bit-Exact ECC Recovery from Write Timing.}
We introduce BREW-RC, the first technique to recover the bit-exact ECC
parity submatrix of a commercial ReRAM product using only write-access
timing, without physical probing or device modification. BREW-RC
formulates ECC recovery as a SAT problem, deriving constraints from
observed write latency distributions and recovering the complete
generator matrix from two commercial products. The recovered matrix
enables accurate timing models in the presence of embedded ECCs and is
validated against results from destructive reverse engineering
techniques.

\paragraph{ECC-Aware Entropy Harvesting.}
We demonstrate that structured variance in commercial ReRAM write
latency is primarily explained by ECC activity and the resulting DHD,
and that prior entropy claims for commercial ReRAM TRNGs are overstated
as a result. Applying ECC-aware analysis, we isolate the deterministic
ECC contribution from the underlying stochastic filament dynamics and
construct stationary, wear-leveled measurement sequences with
min-entropy rigorously estimated under NIST SP 800-90B. We further
characterize how device aging affects switching dynamics and entropy
yield over device lifetime. The resulting TRNG passes NIST SP 800-22 and
PractRand and achieves 17$\times$ throughput improvement over prior
ECC-unaware methods.