\section{Dissertation Organization}
\label{sec:organization} This dissertation is organized into the
following chapters.

\textbf{Chapter~\ref{chap:background}} provides technical background on
the side-channel attack class addressed in this dissertation and the
device behaviors relevant to its analysis. It covers the Pentimenti
class of defect-dynamics-driven side-channels, fine-grained timing
sensor architectures used to extract them, and the device and
architectural behaviors that determine write latency in commercial ReRAM
crossbar memories, including OxRAM bipolar switching dynamics,
program-verify loops, crossbar array architecture, and linear block
codes.

\textbf{Chapter~\ref{chap:fpga-sensor}} closes a vendor-specificity gap
in FPGA Pentimenti measurement. Prior implementations relied on
high-resolution PLL phase shifting available only on Xilinx products.
This chapter introduces a constrained PLL configuration search that
achieves equivalent fine-grained timing resolution on Intel Cyclone V
FPGAs, broadening the platforms on which FPGA Pentimenti analysis is
practical.

\textbf{Chapter~\ref{chap:brewrc}} presents BREW-RC, the first
non-invasive technique to recover the bit-exact ECC parity submatrix
from commercial ReRAM using write-access timing. It describes the
experimental characterization of write latency on two commercial
products, the SAT-based formulation of ECC recovery, and a validation of
the recovered parity matrix against destructive reverse engineering
results.

\textbf{Chapter~\ref{chap:trng}} applies the structural understanding
developed in Chapter~\ref{chap:brewrc} to the characterization of
stochastic write latency in commercial ReRAM. It demonstrates that prior
entropy claims are overstated due to unaccounted ECC activity, presents
an ECC-aware framework for entropy source construction and evaluation,
characterizes the effects of device aging on entropy yield, and
demonstrates a TRNG achieving 17$\times$ throughput improvement over
prior methods.

\textbf{Chapter~\ref{chap:conclusion}} summarizes the contributions of
this dissertation, reflects on the implications for the security of
deployed ReRAM products, and identifies directions for future work.