\section{Entropy Sources and TRNGs}
\label{sec:bg-entropy}

A \emph{noise source} is a physical process that provides
unpredictability. An \emph{entropy source} wraps a noise source together
with digitization and optional conditioning into a subsystem whose
output can seed a cryptographic random bit
generator~\cite{sonmez2016recommendation}. A true random number
generator (TRNG) implements an entropy source in hardware. Unlike
pseudorandom number generators, TRNGs derive randomness from physical
entropy sources whose unpredictability is grounded in device physics
rather than computational assumptions~\cite{sunar2007provably}.
Dedicated entropy source circuits, such as ring oscillator jitter
harvesters~\cite{sunar2007provably} or metastable latch
designs~\cite{tokunaga2008true}, occupy silicon area with no secondary
function. Entropy harvesting repurposes existing on-chip structures as
dual-use entropy sources, amortizing the cost of entropy generation
across components that already serve a storage or compute
function~\cite{holcomb2009power, wang2012flash, kim2019d}.

The per-sample unpredictability of a digitized noise source is
quantified by min-entropy:
\begin{equation}
	\label{eq:hmin}
	H_\infty(X) = -\log_2\!\bigl(\max_{x}\,\Pr[X = x]\bigr),
\end{equation}
which captures the probability of an adversary's best single guess at a
sample value. For an 8-bit symbol alphabet, $H_\infty$ ranges from 0 (deterministic) to 8 (uniform).
Min-entropy is a lower bound on Shannon entropy and is the standard
measure for cryptographic entropy sources because it reflects worst-case
predictability rather than average behavior. A raw noise source with
$H_\infty < 1$ bit per sample is considered insufficient for direct use
as a cryptographic entropy source without conditioning.

A \emph{conditioning} component compresses the raw bitstream according
to the estimated $\hat{H}_\infty$, concentrating source entropy into a
shorter, near-uniform output~\cite{sonmez2016recommendation}. NIST
SP~800-90B distinguishes between \emph{vetted} conditioning functions,
such as SHA-256 and AES-CBC-MAC, whose entropy accounting is performed
analytically under the assumption that the primitive behaves as an ideal
random function, and \emph{non-vetted} functions, whose output must be
evaluated directly by the SP~800-90B estimators. Per the SP~800-90B
vetted conditioning requirements, $\lceil(256 + 64) /
	H_\text{min}\rceil$ input symbols are required per 256-bit output block,
where the 64-bit margin ensures that the conditioning output is
full-entropy with adversarial advantage bounded by $2^{-64}$.

Raw source quality is typically evaluated using the NIST SP~800-90B
estimator suite~\cite{sonmez2016recommendation}, which applies ten
non-parametric min-entropy estimators to a sample sequence of at least
$10^6$ samples and reports the conservative minimum across all
estimators. The suite includes the Most Common Value estimator, the
Collision estimator, the Compression estimator, the t-Tuple estimator,
the LRS estimator, and several prediction-based estimators. The minimum
across all estimators is taken as the conservative $\hat{H}_\infty$ for
the source. A source that fails to produce a meaningful $\hat{H}_\infty$
estimate under SP~800-90B cannot be used as the basis of a cryptographic
TRNG.

Conditioned TRNG output is evaluated using two complementary test
suites. NIST SP~800-22~\cite{rukhin2001statistical} applies fifteen
frequency, serial, spectral, and complexity tests across sub-sequences
of the conditioned bitstream. PractRand~\cite{practrand} is a streaming
battery that evaluates at geometrically increasing data thresholds and
is particularly sensitive to subtle correlations at small sample sizes.
Statistical testing of conditioned output does not constitute an entropy
proof. Instead, it serves as a sanity check that the conditioning
pipeline functions as intended and that no large-scale structure in the
raw source survives conditioning~\cite{sonmez2016recommendation}.

The stochastic nature of resistive switching in ReRAM has motivated its
use as an entropy source. Balatti et al.~\cite{balatti2015true} provided
an early demonstration that cycle-to-cycle SET variability in a single
HfO$_x$ RRAM device can serve as a TRNG. Subsequent work has explored
both resistance-state variability and write-latency variability as
entropy observables in commercial and embedded ReRAM
devices~\cite{suri2018high, pang2019memristors, tolga2024trng}. The
evaluation methodology and results for commercial COTS ReRAM entropy
harvesting are presented in Chapter~\ref{chap:trng}.