\label{chap:trng}

Chapter~\ref{chap:brewrc} demonstrated that embedded ECCs dominate the
observable write latency distributions of commercial ReRAM devices and
that the bit-exact ECC parity submatrix is recoverable through
non-invasive timing analysis. This chapter applies that structural
knowledge to the problem of entropy harvesting from commercial
off-the-shelf (COTS) ReRAM. Prior works on ReRAM-based TRNGs applied
aggressive debiasing to suppress structured variance of unknown origin,
limiting throughput and obscuring the true entropy of the devices. This
chapter shows that the structured variance is fully explained by ECC
activity, that isolating the deterministic ECC contribution from the
underlying stochastic filament dynamics enables rigorous entropy
characterization, and that the resulting TRNG substantially outperforms
prior ECC-unaware approaches. The chapter makes three contributions.
First, it demonstrates that prior entropy claims for commercial ReRAM
TRNGs are overstated because ECC-driven timing perturbations were
conflated with stochastic device behavior. Second, it introduces an
ECC-aware framework for constructing and evaluating stationary,
wear-leveled entropy sources in commercial ReRAM, including
characterization of device aging effects on entropy yield. Third, it
presents a complete TRNG pipeline that passes NIST SP~800-90B,
SP~800-22, and PractRand, achieving 17$\times$ throughput improvement
over prior methods. Section~\ref{sec:trng-structural} presents
experiments confirming that prior structural biases in commercial ReRAM
timing are fully explained by ECC activity.
Section~\ref{sec:trng-optimization} describes the optimization of
bursted, wear-leveled measurement sequences for maximum entropy yield.
Section~\ref{sec:trng-implementation} presents the complete TRNG
implementation pipeline. Section~\ref{sec:trng-results} presents entropy
estimation and statistical validation results.
