\section{Results}
\label{sec:trng-results}

\subsection{NIST SP~800-90B Entropy Estimates}
Table~\ref{tab:nist90b} reports non-IID min-entropy estimates from
$1\mathrm{M}$ samples per source across the nine NIST SP~800-90B
estimators.

\paragraph{4M Entropy Sources.}

All three 4M conditions produce $\hat{H}_{\min} < 0.01$ bits/symbol
(0.0060, 0.0051, and 0.0079 for RT-A, RT-B, and 55\textdegree{}C,
respectively). The compression estimator returns the lowest values
(0.0004 and 0.0046 for RT-A/B), confirming that latency sequences are
highly compressible. All nine SP~800-90B estimators return near-zero
values for every 4M condition, ruling out any estimator-specific
artifact and confirming a fundamental collapse of source entropy under
the runtime measurement regime. The near-zero estimates are consistent
with a latency distribution that has collapsed to a near-deterministic,
narrow support. The compression estimator value of $4\times10^{-4}$
bits/symbol for RT-A approaches the theoretical minimum for a
length-$1\mathrm{M}$ sequence over a 256-bin alphabet. The
elevated-temperature condition yields a marginal improvement to
$\hat{H}_{\min} = 0.0079$ bits/symbol; while thermal activation can
increase filament stochasticity by reducing the energy barrier for
vacancy migration~\cite{yu2012switching}, this effect is insufficient to
rescue the 4M entropy source. The dominant limitation is structural, not
thermal. The runtime entropy collapse is attributed to structural
bursting effects. During characterization, the full sweep of burst
configurations explored the bursting space under a diversity of stimuli
(subsequent bursts are rotated by a single word independent of burst
size). Under the runtime sequence, fixed-width bursts at burst-aligned
offsets within the reserved 256\,B region collapse nearly all write
latency variance. This sensitivity disqualifies the 4M as a viable TRNG
platform under a purely DHD characterization. However, it does motivate
future investigations into why the effects of burst patterns padded with
redundant, non-transitioning words influence entropy.

\input{contents/ICCAD_2026/tables/nist-90b-all-chips}

\paragraph{8M Entropy Sources.}
The 8M device produces $\hat{H}_{\min}$ of 2.4973 (RT-A), 1.6500 (RT-B),
and 1.9844 (55\textdegree{}C) bits/symbol. The $t$-Tuple estimator is
binding across all three conditions (2.8370, 1.7358, 2.2754), consistent
with short-range thermal coupling between adjacent writes. The 0.34-bit
restart penalty on RT-A reflects modest startup bias as the filament
distribution converges after reset. Device-to-device variation of 0.85
bits/symbol between RT-A and RT-B underscores the necessity of per-chip
characterization. The 55\textdegree{}C result suggests thermal
activation provides a real but bounded boost to filament stochasticity.
The contrast with the 4M device is attributed to burst complexity. The
8M crossbar is driven with 32-word bursts while the 4M uses 4-word
bursts. The longer burst sequence aggregates more switching events per
transaction, and while a lower min-entropy is observed than that of the
characterization sequence, the final entropy of the TRNG is
significantly improved over prior methods.

\input{contents/ICCAD_2026/tables/trng-tests-8m}

\subsection{NIST SP~800-22 and PractRand}
Given the near-zero entropy of the 4M under the runtime sequence, it is
excluded from TRNG statistical validation. Table~\ref{tab:nist22}
reports SP~800-22 results for 256\,KB of SHA-256-conditioned output from
all three 8M sources across two independent runs each. All 15 tests pass
for every source-run combination. $p$-values are well-distributed across
$(0,1)$; the lowest observed is 0.01 (Binary Matrix Rank,
55\textdegree{}C-R1). All three streams pass PractRand through 256\,KB
with no anomalies at any evaluation threshold, confirming that no
artifacts from the repeating $S_{opt}$ sequence or wear-leveling pattern
persist through conditioning. The final TRNG achieves a throughput of
1.7\,kbits/s.

\subsection{Relationship to Prior Work}
The 4M failure does not disqualify COTS ReRAM entropy harvesting in
general, nor does it imply that the device contains no stochastic
dynamics. Arul et al.~\cite{tolga2024trng} obtained non-trivial
min-entropy from the same device family by applying randomized write
patterns followed by aggressive von Neumann
debiasing~\cite{von1963various}. The structural latency component
removed in Section~\ref{sec:trng-structural} was on the order of tenths
of a millisecond, while the timer resolution used
in~\cite{tolga2024trng} likely operated at the MHz level and under a Von
Neuman debiasing the least-significant bit of timing measurments likely
reflected true entropy. However, our results indicate that the 4M write
latency is influenced not only by DHD, but also the context of the write
operation. In the characterization experiment, bursts are rotated around
the write buffer in an overlapping manner.

The 8M device performs well in both characterization and runtime
evaluation regimes showing similar entropy in both, and our methods
demonstrate a significant performance improvement over all existing
commercial ReRAM TRNGs.

\subsubsection{Implications for ReRAM Pentimenti}

The entropy of every commercial ReRAM module tested in this chapter was
sensitive to cumulative wear. Because the codeword sequences used for
measurement held DHD fixed, the switching composition presented to each
cell remained constant across the experiment. The only property of the
device that changed was the physical state of its conductive filaments.
The drift in min-entropy observed across cycling therefore reflects
filament degradation propagating through the program-verify loop and out
to the memory interface as a measurable fluctuation in write latency.

This is the same conceptual mechanism that FPGA Pentimenti attacks
exploit in CMOS routing logic. In an FPGA, accumulated BTI stress shifts
transistor threshold voltage. The shift in threshold voltage produces a
shift in propagation delay, which a fine-grained on-chip TDC can
resolve. In commercial ReRAM, cumulative cycling shifts filament
morphology. The shift in morphology changes the number of program-verify
iterations required to complete a write, which the device exposes
through its write-completion handshake at the SPI interface. Both
attacks observe a delay shift that depends on the cumulative bias
history of the underlying device. They differ in the physical mechanism
and in the measurement infrastructure required, but the structure of the
leakage channel is the same.

The TRNG presented in this chapter targeted raw entropy efficiency. Its
codeword sequences were chosen to maximize switching activity per
measurement and to balance wear evenly across the bits of each word. ECC
knowledge supports a wider class of measurement protocols than the TRNG
itself exercises. Without the parity submatrix, no externally applied
write pattern can guarantee DHD properties at the codeword level, and no
balanced wear sequence can be constructed at the bit granularity that
filament defect dynamics operate on. With the parity submatrix, codeword
sequences can be constructed that hold DHD fixed while varying the wear
of individual bit positions. The same construction can concentrate
cycling on a target cell while leaving its neighbors untouched, or it
can isolate SET and RESET cycling histories independently. Future ReRAM
Pentimenti work can evaluate the leakage channel under controlled
defect-dynamics conditions rather than under the worst-case aggregation
that prior entropy-source studies relied on.
