\section{Tunable Dual-Polarity TDC}
\label{sensor_structure}
Our \name TDC\footnote{The sensor architecture and the sensor implemented alongside a PicoRV core executing AES has been open-sourced for the PYNQ-Z2. Additionally, we provide an easy-to-install PYNQ package for interacting with the sensor and an example Jupyter Notebook studying how $\phi$ shifting can be used for isolating relevant computation on the PicoRV running AES. All this is available at: https://github.com/KastnerRG/Tunable-TDC.} has four key features: 1) it captures both rising and falling transition polarities (Dual-Edge); 2) it provides real-time adjustment of the sample window duration; 3) it provides real-time phase adjustment of the sample clock relative to the target computation; 4) it provides real-time frequency adjustment of the sample clock. We use these features to tune the sensor to the voltage fluctuations of the PDN caused by the target.


%Recording both yields more information about the power distribution network than a single transition edge %\rk{is this true? Do we show it has more info? We can comfortably say it behaves differently. Need to make a clear argument here.}.
%2) the \name TDC provides the capability for real-time tuning. Tuning is critical to ensure that we are sampling the PDN at the correct time. The \name TDC uses two clock domains that can be dynamically phase-shifted relative to each other statically or dynamically. This allows fine-tuning the sensor to the device, its environment, and the target computation. It reduces pre-compilation design complexity and avoids timing path violations in tool reports. Bitstream analysis tools use path timing violations to detect TDC sensors~\cite{krautter2019mitigating}. Thus, a sensor without violations is more stealthy because it is less likely to be detected by bitstream analysis tools. 

\input{fig/dl-schema}

Figure~\ref{fig:dl-schema} shows the \name TDC architecture.
The sensor's core is a pulse generator that induces rising and falling transitions through a delay line at a configurable frequency $F_{sample}$.
A single pulse contains a positive (0 $\rightarrow$ 1) and a negative (1 $\rightarrow$ 0) pulse edge.
Positive and negative pulse edges are issued sequentially in the \texttt{Launch Clock} domain.
Pulse edges cause falling and rising transitions to propagate through a linear array of delay line elements to the \texttt{Output} capture register, which is controlled by the \texttt{Capture Clock}.
$\theta$ is the phase difference between the \texttt{Launch Clock} and the \texttt{Capture Clock} -- the time between the pulse launch and the subsequent capture of the transition in the output registers.
When $\theta$ is set correctly, a transition will be propagating through the delay line when the output registers are clocked and record a metastable transition.
The propagation distance is the number delay elements the transition has passed through. 

An example output sequence from two consecutive pulses is shown at the bottom of Figure~\ref{fig:dl-schema}.
Each pulse causes a falling and rising transition to be captured at the output.
Rising Transition 0 shows that the 0 $\rightarrow$ 1 transition reached \texttt{Output[38]}. Falling Transition 0 shows that the 1 $\rightarrow$ 0 propagated to somewhere between \texttt{Output[21]} and \texttt{Output[23]}, with some metastability between the two points.
In the next pulse, Rising Transition 1 propagates differently; the 0 $\rightarrow$ 1 transition propagates to between \texttt{Output[36]} and \texttt{Output[39]}.
Similarly, Falling Transition 1 propagates to between \texttt{Output[20]} and \texttt{Output[23]}.
% Output fluctuations on the rising and falling transitions are highlighted in red in Figure~\ref{fig:dl-schema}.
These changes reflect PDN voltage fluctuations that change the delay line propagation.
The variations provide potential information about the operation of the FPGA, including computation by co-tenants.

The sampling frequency is dictated by the length of the delay line and the speed of the underlying FPGA logic.
If a higher effective sampling frequency is needed, multiple launch/capture clock pairs with a known phase offset can be generated by the clock generator as is done in related work \cite{song20208, choi2020design, wang2016256}.
%This is demonstrated by the grey delay line in the background of Figure~\ref{fig:dl-schema}.
%The remainder of this section describes a sensor with a single delay line.

%\subsection{Power Distribution Network}
%\label{power_distribution_network}
%\dr{Moved from background to here}
%The power distribution network provides voltage to the individual components of an FPGA system. A typical power distribution network consists of a switching voltage regulator that supplies a set of target voltages to the on-board system components.
%The voltage regulator of the power distribution network is designed to provide a target voltage within a design margin.
%With a constant current load, the regulator's output voltage oscillates between high and low values specific to the device's characteristics as internal capacitance is recharged and discharged.
%As the current load in the system increases, the frequency of the oscillations increase.
%Computations in a system consume power, increase the current load, and cause the voltage oscillations to change frequencies.
%The voltage fluctuations caused by an application running on a system can be measured, characterized, and exploited to undermine the security of the system~\cite{zhao2018fpga, schellenberg2018inside}. 
%\cd{TODO worst case use textbf instead of paragraphs}

\paragraph{Pulse Generator:}
The pulse generator produces positive (0 $\rightarrow$ 1) and negative (1 $\rightarrow$ 0) pulse edges that cause falling and rising transitions, respectively, in the delay line.
Each sample produces a rising and a falling transition on the capture registers. A trace is a series of samples.
The pulse generator has two configurable run-time parameters: the sampling frequency $F_{sample}$, which is an integer fraction of the launch clock frequency, and the number of pulses.
Figure~\ref{fig:dl-schema} demonstrates a trace length $=2$, i.e., two rising and two falling transitions. We show that both transitions contain useful information in Section~\ref{results}. This module is implemented in RTL and requires no use of proprietary IP cores for either Intel or Xilinx implementations.

\paragraph{Programmable Clock Generators: }
The \name TDC has two programmable clock generators. On Xilinx architectures, we implement the clock generators using two Xilinx Mixed-Mode Clock Managers (MMCM).
The first MMCM (Figure \ref{fig:dl-schema} \textcircled{\raisebox{-.9pt} {1}}) controls the input clock to the  TDC and the phase relationship $\phi$ between the target clock and the sensor clock. 
%Manipulating $\phi$ allows the sensor to be tuned to optimize the information between the sensor and a co-tenant computation.
Section~\ref{target_clock_tuning} discusses the importance of tuning $\phi$ to capture relevant information about a co-located computation better. A second MMCM (Figure~\ref{fig:dl-schema} \textcircled{\raisebox{-.9pt} {2}} ) generates the launch and capture clocks with a programmable phase offset, $\theta$, between them.
Changing $\theta$ modifies when the pulse generator generates an edge and when the capture clock fires and records the location of the subsequent transition in the output registers. Section~\ref{capture_clock_tuning} demonstrates the importance of tuning $\theta$.

Intel provides programmable clock generation through the PLL Reconfig Intel FPGA IP core. Much like the Xilinx MMCM, this IP core allows for dynamic reconfiguration of all PLL counters and modification to the phase shift of each output counter. Additionally, a set of advanced parameters provides control over the PLL bandwidth, charge pump and loop-filter settings.

During compilation, the TDC sensor is configured to pass timing checks. The phase relationship $\phi$ is unconstrained, and $\theta$ is set to $2\pi$. This means that the TDC sensor cannot be detected by tools that check for timing violations~\cite{krautter2019mitigating}.

\paragraph{Delay Line and Capture Registers: }
\label{delay_line}
For both Xilinx and Intel architectures, the delay line shown in Figure~\ref{fig:dl-schema}, is a series of combinational logic elements that propagate the rising and falling transitions caused by the pulse generator.
The delay elements are constructed from identical digital circuit elements that aim to provide a linear propagation delay, $\tau$.
The delay elements of the TDC should be placed and routed with uniform spacing to ensure consistent delay between each element and a uniform delay through the entire chain.
%Our results show that even a carefully designed delay line exhibits non-ideal behaviors.
%Thus, we argue that tuning the sensor is paramount to accurately extracting information from the PDN.

On Xilinx architectures, the \name TDC uses the fast look-ahead CARRY primitives to create the delay line. The CARRY logic provides a relatively linear delay between each output bit within a single CARRY primitive. 

Intel does not directly provide an IP for constructing carry chains. Instead, carry chains of arbitrary length may be defined through a combination of logic cell (LCELL) primitives configured in arithmetic mode, carry sum buffers, and placement constraints. LCELL primitives are intended to occupy a single adaptive lookup table (ALUT). There are two ALUTs in a single adaptive logic module (ALM). There are 10 ALMs for each logic array block (LAB).  In arithmetic mode, a single ALUT is partitioned into two reduced size LUTs allowing for one ALM to perform two bits of addition with optional pre-adder logic. Therefore, each logic array block can accommodate 20 carry elements. Carry chains must be constructed from vertically contiguous LABs, although carry chains in arithmetic mode on Intel FPGAs can begin at the first or 5th ALM in a LAB allowing carry chains to have a reduced congestion impact. To ensure the sensor delay line maintains a regular structure, we use placement constraints to guarantee the carry chain always starts at the first ALM in each LAB. By default, the Quartus tool will synthesize contiguous carry chains up to a length of 70 carry elements before partitioning them into multiple chains. However, synthesis settings can be modified to allow arbitrary carry chain lengths. In addition to ALUTs, ALMs include four registers, allowing us to pack two capture registers and two delay line outputs into a single ALM \cite{intel_carry-chain}.

The carry logic is configured to compute
\texttt{Output = 65'h0\_ffff\_ffff\_ffff\_ffff + input} so that when \texttt{input} changes from \texttt{65'h0} $\rightarrow$ \texttt{65'h1} on a positive pulse edge, the output of the delay line is a transition with falling polarity from \texttt{Output = 65'h0\_ffff\_ffff\_ffff\_ffff} to \texttt{Output = 65'h1\_0000\_0000\_0000\_0000}.
A transition with rising polarity is produced on the negative pulse edge.


%However, routing delay between CARRY primitives groups introduces non-linear behavior even when the dedicated routing resources are used.
%This means that some output bits are more sensitive to changes in the PDN than others.
%Section~\ref{carray_element_variation} demonstrates the non-linear behavior of the delay line and shows that $\theta$ tuning is critical for maximizing PDN side-channel channel capacity. \cd{forward reference to results comment? } \dr{Nah, we're good. I think I had an awkward page split in my version}

The interplay between the number of bits in the delay line and $\theta$ is also a critical TDC design consideration.
The delay line length limits the maximum value of $\theta$ and the sampling frequency.
A delay line that is too short may not capture all of the PDN variations induced by a target, but a long delay line increases resource consumption.
Characterizing how a target computation affects the PDN and the $\theta$ value that best measures variations are crucial for tuning the sensor to provide the most information.

\input{fig/phi-theta}

The capture registers shown in Figure~\ref{fig:dl-schema} record the output of each bit of the carry delay line in the capture clock domain.
The path from the pulse generator to the high-order bit of the output meets timing constraints in the FPGA toolchain, and the launch clock and capture clock are configured to be in phase during compilation.
This means that the TDC sensor cannot be detected by tools that check for timing violations~\cite{krautter2019mitigating}.

%\cd{Is this paragraph repetitive of the last paragraph in section B?}
%The capture registers record the propagation distance of the rising and falling transitions as shown in Figure~\ref{fig:dl-schema}.
%\dr{First and the last sentence seem redundant and could be condensed}

% In the future, we can say something about sync registers, too
%\dr{Synchronization registers}
% Good point. Let me see how it looks and come back. Or maybe it's not worth referencing in the first place? I want to provide some link in my discussion of the first index to where it is defined

\paragraph{$\theta$ and $\phi$ Tuning: }
The programmable clock generators allow the \name TDC to tune its parameters to optimize the information sampled from the PDN.
Figure~\ref{fig:phi-theta} defines the relationship between the target clock, the capture clock, and the launch clock using $\theta$ and $\phi$, the effect of the target computation on $V_{dd}$, and the effect of varying $\theta$ and $\phi$ on the extracted information.
The PDN voltage $V_{dd}$ varies in response to the target's rising clock edge.

The upper right graph in Figure~\ref{fig:phi-theta} demonstrates the effect of varying $\theta$ from $0$ to $2\pi$. Increasing $\theta$ provides more time for the pulse to propagate through the delay line; as $\theta$ increases, the transition bit index increases. Section~\ref{capture_clock_tuning} experimentally demonstrates the importance of tuning $\theta$.

The lower right graph in Figure~\ref{fig:phi-theta} demonstrates the effect of varying $\phi$ from $-\pi$ to $\pi$. Changing $\phi$ will change the sampling window with respect to the target computation. When the sampling window is correctly positioned to the target clock, the sensor output will maximally change in response to the variations in current drawn by the target. This will cause an increase in the information measured at the sensor. Section~\ref{target_clock_tuning} demonstrates how our TDC sensor enables $\phi$ to be tuned to ensure that the sample window is optimized with respect to the target computation.

\paragraph{Propagation Metric: }
\label{propagation_metric}
When $\theta$ is tuned correctly, the capture clock will record how far the signal has propagated through the delay elements. The signal propagation distance can be measured as the index in the capture register. The least significant bits generally have their post-transition value, and the most significant bits typically have their pre-transition value.
This imprecise definition reflects the metastability around the transition point that can cause multiple bit flips. This metastability may contain useful information, and ignoring these flips could reduce the side-channel information. This behavior is shown in rising/falling Transition 1 of Figure~\ref{fig:dl-schema}.

\input{fig/trace_prop_metrics}

We examine three propagation metrics:
\begin{itemize}
    \item \textit{First Index}: The index of the first bit in \texttt{Output} that is not equal to \texttt{Output[0]}
    \item \textit{Last Index}: The index of the last bit in \texttt{Output} that is equal to \texttt{Output[0]}
    \item \textit{Binary Hamming Distance}: For rising transitions, the binary Hamming distance from \texttt{64'h\_0000\_0000\_0000\_0000}, and for falling transitions, the binary Hamming distance from \texttt{64'h\_ffff\_ffff\_ffff\_ffff}.
\end{itemize}

Figure~\ref{fig:annot_prop_metrics} shows the traces of Figure~\ref{fig:dl-schema} with annotated First index and Last index propagation metrics. The \textit{First Index} metric produces the sequence 39, 21, 37, 20. The \textit{Last Index} sequence is 38, 23, 39, 23. The \textit{Binary Hamming Distance} sequence is 39, 22, 38, 22. Section~\ref{capture_clock_tuning} explores the efficacy of each metric.



\paragraph{Overcoming Coarse Intel Phase Shift Resolution: }
Intel FPGAs provide a minimum phase shift of $\frac{T_{VCOMAX}}{8}$. With the exception of MAX10 devices (96.15ps), all current Intel FPGA product lines have a minimum phase-shift resolution of $78.125ps$ \cite{max10phase, aria10phase, stratix10phase, agilexphase, cvphase}. This is significantly larger than the dynamic interpolated fine phase shift minimum on Xilinx devices of $\frac{T_{VCOMAX}}{56}$. A lower minimum phase shift results in a degradation of sensor resolution for both tuning the observable window ($\theta$) and locating an area of maximum information ($\phi$). To address this limitation, we propose a novel technique to construct a fine-grained phase sweep from lower resolution phase shifts. This technique is described in Algorithm \ref{alg:pll-intel}. 

\begin{algorithm}[t]
\caption{PLL Configuration Prepossessing Algorithm for Intel FPGAs. Given a tuning parameter ($Tune=``phi" or ``theta"$) the TDC frequency ($F_{TDC}$), the target frequency ($F_{TARGET}$), the upstream PLL reference frequency ($F_{ref}$, and FPGA-dependent PLL counter, VCO, and PFD constraints ($Ms$, $Ns$, $Cs$, $F_{vcomin/max}$, $F_{pfdmin/max}$), compute a set of PLL configurations defining a range of phase shift positions offset from either the launch clock (for $Tune=``theta"$) or the PLL initial phase (for $Tune=``phi"$).}\label{alg:pll-intel}

\SetKwInOut{Input}{inputs}
\SetKwInOut{Output}{outputs}
\SetKwProg{ComputePhaseShiftCfgs}{ComputePhaseShiftCfgs}{}{}
\DontPrintSemicolon
\ComputePhaseShiftCfgs{}{
    \Input{$Tune$, $F_{TDC}$, $F_{TARGET}$, $F_{ref}$, $Ms$, $Ns$, $Cs$, $F_{vcomin/max}$, $F_{pfdmin/max}$}
    \Output{Phase-shift-sorted PLL configurations}
    $VCOs \gets Ms \times Ns $ \Comment{Cross multiply Ms and Ns}\;
    $CFGs \gets \emptyset$ \Comment{Data structure to hold valid programmable clk gen. configs}\;
    \If{$Tune == ``theta"$}{%
        $T_{TDC} = \frac{1}{F_{TDC}}$\;
        $Max_{phase} = T_{TDC}$ \;
    } \Else {
        $T_{TARGET} = \frac{1}{F_{TARGET}}$\;
        $T_{TARGET2\pi} = 2 *T_{TARGET} $\;
        $Max_{phase} = T_{TARGET}$ 
    }
    \ForEach{$\left(M, N\right) \in VCOs$}{%
        $F_{vco} \gets \frac{F_{ref}*M}{N}$\; 
        \If{$F_{vcomin} \le F_{vco} \le F_{vcomax}$ }{%
            $F_{pfd} \gets \frac{F_{ref}}{N}$\; 
            \If{ $F_{pfdmin} \le F_{pfd} \le F_{pfdmax}$}{%
                \ForEach{$C \in Cs$}{%
                    $F_{out} \gets  \frac{F_{ref}*M}{N * C}$\;
                    \If{$F_{out} == F_{TDC}$}{%
                        $pb_{delay} \gets \frac{1}{8*F_{vco}}$ \Comment{Time (s) of one minimum phase bump for this PLL config}\;
                        $n_{pbs} \gets \frac{Max_{phase}}{pb_{delay}}$ \Comment{Number of phase bumps in one cycle}\;
                        $PBs \gets \{1, 2, ..., n_{pbs}\}$\;
                        \ForEach{$pb \in PBs$}{ %
                            \If{$Tune == ``theta"$}{%
                                $delay \gets T_{TDC} - pb * pb_{delay}$ \Comment{Delay w.r.t. the launch clock}\;
                            } \Else {
                                $delay \gets pb * pb_{delay}$ \Comment{Delay w.r.t. to zero phase shift}\;
                            }
                            $CFGs_{\theta}.append(\{M, N, C, pb, delay\})$\;
                        }
                    }
                }
            }
        }
    }
    $CFGs \gets sort(CFGs)$ \Comment{Sort configs w.r.t. counter configurations M, N, C, phase (s)}\;
    $CFGs \gets drop\_duplicates(CFGs, delay)$ \Comment{Remove identical configurations w.r.t phase delay}\;
}
\end{algorithm}

To improve the minimum phase shift for $\theta$ and $\phi$, we construct every possible phase shift increment by exhaustively iterating through the space of all possible valid PLL configurations. A valid PLL configuration must conform to the specifications provided by the part manufacturer. For Cyclone V, the set of valid configurations must have M, N, and C counters in the range of 1 to 512, a $F_{vco}$ in the range of 600-1600MHz, and a $F_{pfd}$ in the range of 5-325MHz \cite{intel_pll-spec}. From these valid configurations we determine all possible phase shifts either with respect to the launch clock or the zero phase shifted position for $\theta$, and $\phi$, respectively. This results in a much more fine grained (although variable) phase shift resolution. Each PLL configuration will contribute varying power consumption, phase noise and jitter due to the modification of the PLL counters, VCO gain, loop filter and charge pump settings. Our results indicate that the statistical methods used to tune $\theta$ and $\phi$ accommodate these perturbations, allowing us to adjust the position of the observable window and identify an area of maximum information despite the noise contributed by PLL reconfiguration. We note that there is no additional target knowledge or a change of threat model to use Algorithm \ref{alg:pll-intel}.

\input{fig/Intel-phase-dist}

To maximize the sampling rate of the \sensor TDC, the programmable clock generator \textcircled{\raisebox{-.9pt} {2}} should operate at a maximum frequency. However, the upstream programmable clock generator \textcircled{\raisebox{-.9pt} {1}} can operate at many frequencies without impact to the sampling rate. Fine adjustment is more critical for $\phi$, as we do not have control over the phase or frequency of the target co-tenant logic. Figure \ref{fig:phase-shift-intel} shows the distribution of minimum phase increments for both $\theta$ and $\phi$ for a fixed upstream clock generator frequency of 10MHz and fixed downstream frequency of 100MHz. The minimum PLL phase shift of $\frac{T_{VCOMAX}}{8}=78.125ps$ becomes a worst-case condition for both tuning parameters. However, on average, our minimum phase shift improves for both $\theta$ and $\phi$ to 1.6ps and 15.527ps, respectively. The upstream clock generator can operate at a lower frequency, yielding more PLL arrangements for tuning $\phi$. The downstream clock generator must provide a clock signal to the TDC logic at 4 times the sensor sampling rate. This restriction presents a trade-off between the sampling rate and the resolution of $\theta$. As the PLL frequency scaling factor $\frac{F_{out}}{F_{ref}}$ is increased given a fixed $F_{ref}$ there are fewer integer solutions to M, N, and O in the equation $\frac{F_{out}}{F_{ref}}=\frac{M}{N*O}$. For example, in our demonstrated PLL, half as many integer solutions exist to the equation $\frac{100MHz}{10MHz} =\frac{M}{N*O}$ for M, N, and O bounded to the range (1, 512) than there would be for $\frac{50MHz}{10MHz} =\frac{M}{N*O}$. Despite this trade-off, our phase shifting technique for $\theta$ still yields a significantly improved average phase step resolution at a launch and capture clock rate of 100MHz (25MHz TDC sampling rate). 

Unique integer PLL configurations that produce identical output frequencies do not produce equivalent phase noise, jitter, and power consumption. However, with a sufficiently large sample sizes, we observe in Sections \ref{capture_clock_tuning} and \ref{target_clock_tuning} that these noise sources do not impair our ability to tune $\phi$ and $\theta$. Additionally, modifying integer PLL counters requires a reset to the reconfigured PLL. This is not an issue for tuning $\theta$, as the phase delay is measured between launch and capture clocks. However, $\phi$ is measured from the zero phase shifted initial position of the PLL and resetting the PLL can perturb $\phi$ tuning as resetting Intel PLLs will erase all prior phase shift settings and results in a loss of PLL lock. We describe this limitation in greater detail in Section \ref{target_clock_tuning} and provide a solution that resets the PLL only at the start of each consecutive coarse-grained $\phi$ sweep.

% To reduce phase noise and jitter at the output of the PLLs we identify equivalent PLL configurations that yield more desirable stability characteristics. In Algorithm \ref{alg:pll-intel}, we implement a drop\_duplicates function that identifies the most stable PLL outputs, and drops other configurations with matching phase offsets. To improve the quality of PLL configurations further, we specify a phase step resolution ($t_{PS}$) where the set of PLL configurations in the range ($i*(\frac{t_{PS}}{2} to (i+1)*\frac{t_{PS}}{2}$) is sorted for the best phase noise and jitter performance. To compare the DE10-Nano against it's closest Xilinx counterpart (PYNQ-Z2), we set $t_{PS}$ to 14ps. For integer PLLs, jitter can be improved by increasing the loop filter capacitance, $F_{PFD}$, and VCO gain.

%The \textit{First Index} and \textit{Last Index} metric work well if there is a single transition or if bits in the metastable region are not correlated with PDN variations and the inter-CARRY primitive delay is small. In contrast, the \textit{Binary Hamming Distance} metric counts multiple transitions and can increase side-channel information if the metastability is correlated with PDN variations. 


%The \textit{First Index} and \textit{Last Index} metric work well if there is a single transition or if bits in the metastable region are not correlated with PDN variations and the inter-CARRY primitive delay is small. In contrast, the \textit{Binary Hamming Distance} metric counts multiple transitions and can increase side channel information if the metastability is correlated with PDN variations.
%In the remainder of this paper, we show that the \textit{Binary Hamming Distance} metric optimizes the channel information because there is frequently more than one apparent transition in a sample and inter-primitive timing delay is not linear.
