\input{fig/propagation}
\input{fig/prop_metrics_Intel}

\subsection{$\theta$ Tuning and Metric Selection}
\label{capture_clock_tuning}

As shown in Figure~\ref{fig:phi-theta}, $\theta$ is the phase difference between the launch and capture clocks and dictates how long a transition is allowed to propagate through the delay line. It plays two important roles: first, $\theta$ determines the position of the transition in the output and can be used to avoid undesirable behavior caused by discontinuities in the FPGA architecture; second, $\theta$ defines the duration of the sampling window, when the delay line is measuring PDN variations.

Figure~\ref{fig:propagation} and Figure~\ref{fig:propagation-intel} demonstrate the effect varying $\theta$ has on the transition index as measured by the \textit{First Index}, \textit{Last Index}, and \textit{Binary Hamming Distance} metrics across both falling and rising transition polarities for Xilinx and Intel, respectively.
These experiments are performed on the {\it PYNQ-Z2 Sensor Only}, {\it AWS Sensor Only} and {\it Intel DE10-Nano} Sensor Only designs. In the Xilinx experiment $\theta$ is increased from 0~ps with a step size of 11.16~ps on AWS and 14.88~ps on PYNQ, as determined by the maximum $F_{vco}$ frequency for the family and device speed grade. In the Intel experiment we iterate through the constructed set of phase delays for $\theta$ corresponding to the bottom phase-step distribution shown in Figure~\ref{fig:phase-shift-intel}. For both platforms at each value of $\theta$ a trace of $2^{14}$ samples is captured, where a sample is one rising and one falling transition. This process is repeated on Xilinx until the transition index exceeds 64 bits, the maximum length of the delay line for our PYNQ-Z2 implementation. for Intel we sweep the full length of the 128-bit TDC. Next, for both architectures, we calculate the transition index using \textit{First Index}, \textit{Last Index}, and \textit{Binary Hamming Distance} metrics, for each value of $\theta$, for each trace, for both rising/falling transition polarities.

The average value of the trace at each value of $\theta$ is plotted. Expressed as the error bar at each point is the standard deviation of the respective trace. Standard deviation, as we will show, is a good measure of the sensitivity of the sensor to voltage changes. In the Xilinx plots, rising/falling transition polarities are shown in blue/orange for AWS and red/green for PYNQ. For the DE10-Nano, rising/falling transition polarities are shown in red/green. The three sub-graphs correspond to the three propagation metrics from Section~\ref{propagation_metric} which are studied in the following sections. As a point of comparison, the black dots annotated onto each plot represent the worst-case minimum phase-shift on Intel FPGAs. All inferences made from trace features visible between these points would not be observable on Intel FPGAs without the use of our PLL Configuration Prepossessing Algorithm.

%Figure~\ref{fig:prop-first} demonstrates the behavior of the rising and falling transitions as measured by the \textit{First Index} metric on both PYNQ-Z2 and AWS. Irregularities in the propagation of the rising transition are immediately apparent within both architectures, where sloped regions reflect the 

\paragraph{First Index: }
\label{first_transition}
From Figure~\ref{fig:prop-first} we see that irregularities in the propagation of the rising transition are immediately apparent within both the PYNQ-Z2 and AWS. Plateaus where the transition index does not increase appear for ${\sim}40$~ps (${\sim}5$~ps) on PYNQ-Z2 (AWS) are divided by sloped regions where propagation is significantly faster at ${\sim}.15\frac{bits}{ps}$ (${\sim}.5\frac{bits}{ps}$) on PYNQ-Z2 (AWS). The sloped regions span four (eight) bits on PYNQ-Z2 (AWS) reflecting the underlying CARRY4 (CARRY8) primitives. Figure~\ref{fig:prop-first-intel} demonstrates that the first transition propagation metric for the DE10-Nano displays non-linearities that do not directly map to carry elements. The propagated carry signal enters the chain at the first carry element of a LAB and every 20 output bits correspond to a new LAB. Deviations in linearity are observable but do not correspond to any particular placement irregularity or grouping of carry chain logic. Unlike the Xilinx implementation where the delay line is composed of CARRY4/CARRY8 primitives, the Intel implementation is composed of individual carry elements and unsynthesized carry buffers that instruct the compiler to synthesize a contiguous chain. The underlying physical circuitry is not known. Therefore, these deviations may be due to the underlying physical logic or may be a byproduct of changes in the clock generator \textcircled{2} PLL driving the capture registers. The 128-bit DE10-Nano delay line propagates transitions at a rate of of $.07\frac{bits}{ps}$ and in the sloped region propagates at a rate of $.15\frac{bits}{ps}$. Low variance is observed in most of the measurement period, with the rising edge showing elevated variance in three segments of the sloped region, and the falling edge showing little variance at all.

%The falling transition notably does not produce such pronounced plateau/slope artifacts.
%These sloped regions show the greatest standard deviation at each point, in contrast to the plateaus which show little to none.
%Figure~\ref{fig:prop-first} demonstrates the behavior of the rising and falling transitions as measured by the \textit{First Index} metric on both PYNQ-Z2 and AWS. Irregularities in the propagation of the rising transition are immediately apparent within both architectures. Plateaus, regions where the transition index does not increase, appear for ${\sim}40$~ps on Zynq and ${\sim}5$~ps on AWS/UltraScale+. The plateaus are interleaved with sloped regions where we see propagation is significantly faster on the UltraScale+ part (${\sim}.5\frac{bits}{ps}$) than on Zynq (${\sim}.15\frac{bits}{ps}$). The sloped regions also reveal the differences in underlying structure: the sloped regions span four bits on Zynq, reflecting the CARRY4 primitives, and span eight bits on AWS/UltraScale+, reflecting the CARRY8 primitives. These sloped regions show the greatest standard deviation at each point, in contrast to the plateaus which show little to none.

%Between the rising and falling transitions, the falling transition does not produce such pronounced plateau/slope artifacts. Its overall propagation is more consistent with fewer and less significant plateaus compared to the rising transition. In addition there are fewer points with zero standard deviations.

\paragraph{Last Index: }
\label{last_transition}
Figure~\ref{fig:prop-last} and Figure~\ref{fig:prop-last-intel} demonstrates the behavior of the rising and falling transitions as measured by the \textit{Last Index} metric on both PYNQ-Z2/AWS and Intel, respectively. As demonstrated, plateaus are still present on both architectures but they are less prominent and the standard deviation is more consistent across values of $\theta$ on the line. On the Xilinx systems, noticeable plateaus still exist where the rising transition resembles that of the \textit{First Index}. Plateaus are obvious on the AWS device and less on the PYNQ-Z2. On the DE10-Nano, deviations in the slope of are still present and still do not map to known physical delay line structures. However, variance is significantly improved on the DE10-Nano using the last index propagation metric.

%Again, the falling transition demonstrates fewer plateaus and more consistent standard deviation at each point.

%As with the \textit{First Index} metric, between the rising and the falling transition, the falling transition demonstrates fewer plateaus and more consistent standard deviation at each point. The difference between the rising and falling transitions is significantly less with the \textit{Last Index} metric.


\paragraph{Binary Hamming Distance: }
Figure~\ref{fig:prop-pop} and Figure~\ref{fig:prop-pop-intel} demonstrates the behavior of the rising and falling transitions as measured by the \textit{Binary Hamming Distance} metric on PYNQ-Z2/AWS, and the DE10-Nano, respectively. On the Xilinx systems, no prior method accounts for metastability within TDC output. For example, if a rising transition falls within \texttt{Output[36:40]} as in Figure~\ref{fig:dl-schema}, neither prior metric is able to discern between \texttt{4b'0101} and \texttt{4b'0111}, potentially missing important information. The data in Figure~\ref{fig:prop-pop} demonstrates that there are few plateaus when using the \textit{Binary Hamming Distance} metric, and that the standard deviation is relatively consistent across the delay line. On the DE10-Nano, only a very small plateau is observed in prior metrics in the sloped region at 1445ps and and the position of the transition monotonically increases for the rest of the measurement. However at lower sample sizes the average transition points vary dramatically. Hamming weight acts as a more linear propagation metric across the measurement period.   

The variable $\theta$ provides the ability to choose where in the delay line a transition falls, and therefore the ability to avoid plateaus we have observed in this section. For the remainder of the paper we use the \textit{Binary Hamming Distance} metric for measuring rising and falling polarities due to its improved characteristics.

On the Xilinx systems, the delays of the carry outputs do not monotonically increase due to the use of carry lookahead adders in the FPGA architecture. Permuting the outputs allows the timing to be maintained~\cite{glamovcanin2020cloud}. This would change the behavior of the first/last index metric, making them more linear. It does not effect the \textit{Binary Hamming Distance}. On both architecture, using the hamming weight metric provides a more linear, consistent delay line measurement.

%\vspace{-10pt}
%The data in Figure~\ref{fig:prop-pop} demonstrates that there are few plateaus when using the \textit{Binary Hamming Distance} metric, and that the standard deviation is relatively consistent across the delay line. As the case in the previous two metrics, the falling transition is less affected by CARRY primitive boundaries and has a more consistent standard deviation across all points. However, the maximum standard deviation is less than that of the first index metric.

%The data in Figure~\ref{fig:prop-pop} demonstrates that there are few plateaus when using the \textit{Binary Hamming Distance} metric, and that the standard deviation is relatively consistent across the delay line. As the case in the previous two metrics, the falling transition is less affected by CARRY primitive boundaries and has a more consistent standard deviation across all points. However, the maximum standard deviation is less than that of the first index metric.

%\paragraph{Analysis:}
%A careful choice in metric of propagation and transition polarity is clearly important because it can affect the sensitivity and range of the sensor. 
%Circuits can cause characteristic amounts of PDN variation, and some circuits may cause small variations and some circuits may cause large variations. 


%For example, if \textit{First Index} and positive transition polarity is chosen as a metric of propagation, it is possible to tune $\theta$ such that the transition lands in a region with high standard deviation. This may be extremely sensitive, but this offers at most 4 bits of swing on PYNQ-Z2 and 8 bits of swing on UltraScale+. Outside of those regions disruptions in the PDN may be lost in the surrounding plateaus. %A consistent slope and variance at each point is desirable because it leads to high sensitivity and predictable response in the presence of large and small PDN variations. This is useful in our attack scenario where we are trying to determine the type of core and assume nothing about the variations on the PDN.

%For the remainder of the paper we use the \textit{Binary Hamming Distance} metric for measuring rising and falling polarities. This selection reduces the number of plateaus in comparison to the other combinations (\textit{First Index} rising, \textit{First Index} falling, \textit{Last Index} rising, \textit{Last Index} falling). It also provides consistent standard deviation compared with these  combinations, which demonstrates sensitivity at each point.