\subsection{Multi-tenant effects on Tuning}

In a shared multi-tenant FPGA cloud infrastructure, more than one additional tenant may occupy and vacate an FPGA resource while an adversary performs an attack. Capture clock tuning is dependant on the relative variance of different positions on the delay line. This tuning procedure can be carried out independent of tenant occupancy. However, target clock tuning, background subtraction, and side-channel attacks may be impacted by additional tenants entering and exiting the shared FPGA resource. We now consider multi-tenant effects on our proposed tuning techniques.

\paragraph{Target tuning and side-channel analysis: } During target tuning, arriving (exiting) tenants would increase (reduce) variance in the sensor measurements. If an arriving victim computation is not phase-aligned to the adversary, the adversary would observe additional peaks in variance at the frequency and phase of the new tenant logic. Figures \ref{fig:phi-intel} and \ref{fig:phi_xilinx} demonstrate this scenario as the out of phase shell logic (AWS) or processor interface logic (Pynq Z2 and DE10-Nano) is clearly distinguishable in the measured $\phi$ sweep. If the arriving tenant is phase-aligned to the target tenant, the adversary would observe a stronger peak in variance, enhancing the identification of the target phase. However, multiple phase aligned victims would act as noise sources for subsequent side-channel attacks, decreasing attack efficiency. In this scenario, the activity from additional (non-target) tenants could be described as a power wasting circuit, which have been demonstrated to reduce the signal-to-noise ratio of trace measurements \cite{krautter2019active}.

\paragraph{Background subtraction: } Our target tuning and background subtraction techniques require approximately one second on AWS and a few minutes on the Pynq Z2 and DE10-Nano platforms to measure background noise. During a sweep of $\phi$ to collect background traces, tenants arriving or vacating the shared FPGA resource may perturb measurements. In this context, the adversary may under/overcompensate for background noise. To detect a change in occupancy, an adversary could measure contention of a shared resource \cite{giechaskiel2022cross}. With this capability, an adversary could wait for a period of stable occupancy to collect background traces or subtract the perturbations induced by new tenants.
